Ore Polynomials in Sage

### 2013-06-16 paper/main.tex
paper: minor changes

At the time of writing, the package we describe here is still under construction
and has not yet been incorporated into the official Sage distribution. Readers
who want to try it out are invited to download the current latest version from 
\begin{center}
  URL
\end{center}
and are encouraged to send us bug reports and other comments. We hope 
that the community will find the code useful.

Creating an Ore algebra object is easy. The following instructions show how to 
load the code and then create an Ore algebra $A$ of linear differential operators
and an Ore algebra $B$ of recurrence operators. Observe the correct application
of the respective commutation rules in both cases.

@@

The package also supports Ore algebras with several generators already. However,
so far we offer no functionality beyond addition and multiplication for
these. Much more functionality is available for algebras with one generator. We
describe some of it in Section~\ref{sec:4}.

@@

An Ore algebra is determined by a base ring and a finite number of generators.
In the examples above, the base ring was $\Bold{Z}[x]$, and the generators were
$\mathrm{Dx}$ and $\mathrm{Sx}$, respectively. If no other information is provided
in the arguments, the \verb|OreAlgebra| constructor chooses the nature of the 
generators according to their name: a generator called $\mathrm{Dt}$ represents
the standard derivation $d/dt$ acting on the generator $t$ of the base ring, 
a generator called $Sn$ represents the standard shift operator sending the
generator $n$ of the base ring to $n+1$.

@@

In general, the commutation rules of a generator~$X$ of an Ore algebra~$A$ with
base ring~$R$ are governed by two maps, $\sigma\colon R\to R$ and $\delta\colon R\to R$,
where $\sigma$ is a ring endomorphism (i.e., $\sigma(a+b)=\sigma(a)+\sigma(b)$ and
$\sigma(ab)=\sigma(a)\sigma(b)$ for all $a,b\in R$) and $\delta$ is a skew-derivation
for $\sigma$ (i.e., $\delta(a+b)=\delta(a)+\delta(b)$ and $\delta(ab)=\delta(a)b
+\sigma(a)\delta(b)$ for all $a,b\in R$). With two such maps being given, the
generator $X$ satisfies the commutation rule $Xa=\sigma(a)X+\delta(a)$ for every
$a\in R$. If there is more than one generator, then each of them has its own pair
of maps $\sigma,\delta$. Different generators commute with each other; 
noncommutativity only takes place ``between'' generators and base ring elements.

@@

In the dictionary specifying~$\sigma$, omitted generators are understood to be
mapped to themselves, so that \verb|{}| in the definition of $A$ in the example
above is equivalent to \verb|{x:x,y:y,z:z}|.  Likewise, in the dictionaries
specifying~$\delta$, omitted generators are understood to be mapped to zero.

For Ore algebras with several generators, it is possible to mix specifications
of generators via triples $(X,\sigma,\delta)$ with generators using the naming
convention shortcuts as explained before. Continuing the previous example,
here is a way to define an algebra $A$ over $K$ with two generators, a $D$ 
that behaves like before, and in addition an $Sx$ which acts like the standard
shift on~$x$.

\begin{sageexample}
  sage: A = OreAlgebra(K, ('D', {}, {x:1, y:y, z:1/x}), 'Sx')
  sage: D, Sx = A.gens()
  sage: D*x, Sx*x
\end{sageexample}

@@

Besides basic arithmetic for Ore operators, the package provides a wide range of
methods to create, manipulate and solve several different kinds of operators.
Some of these methods are accessible in any Ore algebra while others are tied
specifically to e.g.\ recurrence operators or differential operators.

In this section, we give an overview of some of the functionality provided by
the package. Because of space limitation, only some of the available methods
can be discussed here. For further information, we refer to the documentation.

@@

\begin{center}
  \begin{tabular}{|l|p{.65\hsize}|}
    \hline
    Method name & short description \\\hline
    \verb|A.quo_rem(B)| & returns the left quotient and the left remainder of $A$ and $B$.\\
    \verb|A.gcrd(B)| & returns the greatest common right divisor of $A$ and $B$.\\
    \verb|A.xgcrd(B)| & returns the greatest common right divisor of $A$ and $B$ and the according B\'ezout coefficients.\\
    \verb|A.lclm(B)| & returns the least common left multiple of $A$ and $B$.\\
    \verb|A.xlclm(B)| & returns the least common left multiple $L$ of $A$ and $B$ and the left quotients of $L$ and $A$ and of $L$ and $B$.\\
    \verb|A.resultant(B)| & returns the resultant of $A$ and $B$ (see \cite{li96} for further information).\\
    \hline
  \end{tabular}
\end{center}

@@

The GCRD is only unique up to multiplication (from the left) by an element from
the base ring. The method \verb|normalize| called in line~6 of the listing above
multiplies a given operator from the left by some element from the base ring
such as to produce a canonical representative of the set of all the operators
that can be obtained from the given one by left multiplication of a base ring
element. This facilitates the comparison of output.

There are a number of options to adjust the functionality of the methods. For
example, the efficiency of computing the GCRD depends on the size of the
coefficients of intermediate results and there are different strategies to
control this in so called polynomial remainder sequences (PRS). The default is
the improved PRS described in~\cite{jaroschek13b}. which will usually be the
fastest choice.

@@

If $L_1,L_2$ are operators, then the solutions of their GCRD are precisely the
common solutions of $L_1$ and~$L_2$. The LCLM, on the other hand, is the minimal
order operator whose solution space contains all the solutions of $L_1$ as well
as all the solutions of~$L_1$. Because of this property, GCRD and LCLM are
useful tools for constructing operators with prescribed solutions. For example,
here is how to construct a differential operator which has the solutions $x^5$
and $\exp(x)$ starting from the obvious operators annihilating $x^5$
and~$\exp(x)$, respectively.

@@

Thanks to the LCLM operation discussed above, we have the property that when $f$
and $g$ are two objects which are annihilated by some operators $L_1,L_2$
belonging to some Ore algebra~$A$ then this algebra contains also an operator
which annihilates their sum~$f+g$. Thus the class of solutions of operators of
$A$ is closed under addition. It turns out that similar closure properties hold
for other operations. The following table lists some of the corresponding
methods. Methods for some further closure properties will appear further below.

@@

As an example application, let us prove Cassini's identity for Fibonacci
numbers: 
\[
  F_{n+1}^2 - F_nF_{n+2}=(-1)^n.
\]
The idea is to derive, using commands
from the table above, a recurrence satisfied by the left hand side and then show
that this recurrence is also valid for the right hand side.

@@

For the elements of some of the most important algebras, additional methods are
available. The following table lists some of the additional methods available
for differential operators, i.e., elements of an Ore algebra of the
form~$R[x]\<Dx>$ or $K(x)\<Dx>$.

@@

The last output implies that the operator annihilating $\int_0^x \exp(-t^2) dt$
admits also a solution which behaves for $x\to\infty$ like $\frac1x\exp(-x^2)$.

The next example illustrates the methods for finding rational and polynomial
solutions of an operator~$L$.  These methods accept as an optional parameter an
inhomogeneous part consisting of a list (or tuple) of base ring elements,
$(f_1,\dots,f_r)$. They return as output a list of tuples $(g,c_1,\dots,c_r)$
with $L(g)=c_1f_1+\cdots+c_rf_r$ where $g$ is a polynomial or rational function
and $c_1,\dots,c_r$ are constants, i.e., elements of the base ring's base
ring. The tuples form a basis of the solution space.

In the example session below, we start from two polynomials $p,q$, then compute an 
operator $L$ having $p$ and $q$ as solutions, and then recover $p$ and $q$ from~$L$.
Note that for consistency also the solutions of homogeneous equations are returned as 
tuples. At the end we give an example for solving an inhomogeneous equation.

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values of a sequence can be
computed asymptotically faster for large $n$ than by listing all
values, using the binary splitting technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $e$ takes a fraction of a second:

@@

Guessing is, in some sense, the reverse operation of \verb|to_list| for
recurrence operators, or \verb|power_series_solutions| for differential
operators.  It is of the most popular features of packages like gfun, and there
are even some special purpose packages dedicated to this
technique~\cite{kauers09a,hebisch11}. The basic idea is simple. If we are given
a finite array of numbers, thought of as the first terms of an infinite
sequence, we may want to know whether this sequence satisfies a recurrence. The
algorithm behind a guessing engine searches for small equations matching the
given data.  Generically, no such equations exist, so if some are found, it is
fair to ``guess'' that they are in fact valid equations for the whole infinite
sequence.

@@

A number of options are available for governing the search for relations.  In
order to explain them, we first need to give some details on the underlying
algorithms. For simplicity of language, we restrict here to the case of
recurrence operators. The situation for differential operators is very similar.

@@

\hangindent=-6.3cm\hangafter=-13\leavevmode
\smash{\raise-5.3cm\rlap{\kern\hsize\kern-5.1cm\begin{tikzpicture}[scale=.45]
  \draw[->] (10.5,0) node {$r$} (-.2,0) -- (10, 0) ;
  \draw[->] (0,10.5) node {$d$} (0,-.2) -- (0, 10) ;
  \draw (1.2, 3.5) -- (1.4, 4) (1.2, 3) -- (1.6, 4) (1.2, 2.5) -- (1.8, 4) (1.2, 2) -- (2, 4)
        (1.4, 2) -- (2.2, 4) (1.6, 2) -- (2.4, 4) (1.8, 2) -- (2.6, 4) (2, 2) -- (2.8, 4) 
        (2.2, 2) -- (3, 4) (2.4, 2) -- (3.2, 4) (2.6, 2) -- (3.4, 4) (2.8,2)--(3.6,4) (3,2)--(3.8,4)
        (3.2, 2) -- (4, 4) (3.4, 2) -- (4.2, 4) (3.6, 2) -- (4.4, 4) (3.8,2)--(4.6,4) (4,2)--(4.8,4)
        (4.2, 2) -- (5, 4) (4.4, 2) -- (5.2, 4) (4.6, 2) -- (5.4, 4) (4.8,2)--(5.6,4) (5,2)--(5.8,4); 
  \fill[lightgray] (.75, 10) .. controls (1.5,2.5) and (2.5,1.5) .. (10, .75) -- (10, 10) -- (.75, 10);
  \draw[thick] (2, 10) .. controls (2, 1.2) and (2, 1.2) .. (10, 1.2); 
  \draw (-.2, 2) node {\vbox{\llap{\footnotesize min }\kern-3pt\llap{\footnotesize degree }}} -- (10, 2) 
        (-.2, 4) node {\vbox{\llap{\footnotesize max }\kern-3pt\llap{\footnotesize degree }}} -- (10, 4);
  \draw (1.2, -.2) node {\hbox to0pt{\rule{0pt}{2em}\hss\footnotesize min order\hss}} -- (1.2, 10)
        (5, -.2) node {\hbox to0pt{\rule{0pt}{2em}\hss\footnotesize max order\hss}} -- (5, 10);
  \fill (2.5, 3) circle(4pt) node[right] {\rule[-1em]{0pt}{0pt}$A$} (2, 9) circle(4pt) node[right] {$B$} ;
\end{tikzpicture}}}%
The figure on the right illustrates the typical situation for guessing problems
that are not too small and not too artificial. The light gray region indicates
the area which is not accessible with the given amount of data. Only the points
$(r,d)$ below it can be tested for an operator of order~$r$ and degree~$d$ that
fits to the given data. Let's assume that operators exist on and above the solid
black hyperbola. The user will usually not know this curve in advance but may
have some expectations about it and can restrict the search accordingly, for
example to the dashed area shown in the figure. The method will detect the
existence of an operator, say at point~$A$, and construct from the information
gained at this point an operator of minimal possible order, which may correspond
to point~$B$.  This operator is returned as output. Note that the degree of the
output may exceed the value of \verb|max_degree|, and its order may be smaller
than $\verb|min_order|$:

@@

In order to test a specific point $(r,d)$, the data array must contain at least $(r+1)(d+2)$ terms. 
If it has more terms, the guess becomes more trustworthy, but also the computation time increases.
By setting the option \verb|ensure| to a positive integer~$e$, the user can request that only 
such points $(r,d)$ should be tested for which the data array contains at least $e$ more terms than
needed. This increases the reliability. By setting the option \verb|cut| to a positive integer~$c$,
the user requests that for testing a point $(r,d)$, the method should take into account at most $c$
more terms than needed. If the data array contains more terms, superfluous ones are ignored in the
interested of better performance. We must always have $0\leq e\leq c\leq \infty$. The default setting
is $e=0$, $c=\infty$.

>>>>

At the time of writing, the package we describe here is still under construction
and has not yet been incorporated into the official Sage distribution. Readers
who want to try it out are invited to download the current version from 
\begin{center}
  \verb|http://www.risc.jku.at/research/combinat/software/ore_algebra|
\end{center}
and are encouraged to send us bug reports or other comments. We hope 
that the community will find the code useful.

The following instructions show how to load the code and then create an Ore
algebra~$A$ of linear differential operators and an Ore algebra~$B$ of
recurrence operators. Observe the correct application of the respective
commutation rules in both cases.

@@

The package also supports Ore algebras with several generators already. However,
so far we offer hardly more functionality than addition and multiplication for
these. Much more functionality is available for algebras with one generator. 
Some of it is described in Section~\ref{sec:4}. We plan to add more code for the
multivariate case in the future.

@@

An Ore algebra is determined by a base ring and a finite number of generators.
In the examples above, the base ring was $\Bold{Z}[x]$, and the generators were
$\mathrm{Dx}$ and $\mathrm{Sx}$, respectively. If no other information is provided
in the arguments, the \verb|OreAlgebra| constructor chooses the nature of the 
generators according to their name: a generator called $\mathrm{Dt}$ represents
the standard derivation $d/dt$ acting on the generator $t$ of the base ring, 
a generator called $\mathrm{Sn}$ represents the standard shift operator sending the
generator $n$ of the base ring to $n+1$.

@@

In general, the commutation rules of a generator~$X$ of an Ore algebra~$A$ with
base ring~$R$ are governed by two maps, $\sigma\colon R\to R$ and $\delta\colon R\to R$,
where $\sigma$ is a ring endomorphism (i.e., $\sigma(a+b)=\sigma(a)+\sigma(b)$ and
$\sigma(ab)=\sigma(a)\sigma(b)$ for all $a,b\in R$) and $\delta$ is a skew-derivation
for $\sigma$ (i.e., $\delta(a+b)=\delta(a)+\delta(b)$ and $\delta(ab)=\delta(a)b
+\sigma(a)\delta(b)$ for all $a,b\in R$). With two such maps being given, the
generator $X$ satisfies the commutation rule $Xa=\sigma(a)X+\delta(a)$ for every
$a\in R$. If there is more than one generator, then each of them has its own pair
of maps $\sigma,\delta$. Different generators commute with each other; 
noncommutativity only takes place between generators and base ring elements.

@@

In the dictionary specifying~$\sigma$, omitted generators are understood to be
mapped to themselves, so that \verb|{}| in the definition of $A$ in the example
above is equivalent to \verb|{x:x,y:y,z:z}|. In the dictionaries
specifying~$\delta$, omitted generators are understood to be mapped to zero.

For Ore algebras with several generators, it is possible to mix specifications
of generators via triples $(X,\sigma,\delta)$ with generators using the naming
convention shortcuts as explained before. Continuing the previous example,
here is a way to define an algebra $A$ over $K$ with two generators, a $D$ 
that behaves like before, and in addition an $Sx$ which acts like the standard
shift on~$x$ and leaves the other generators fixed.

\begin{sageexample}
  sage: A = OreAlgebra(K, ('D', {}, {x:1, y:y, z:1/x}), 'Sx')
  sage: D, Sx = A.gens()
  sage: D*x, Sx*x
  sage: D*y, Sx*y
  sage: D*z, Sx*z
\end{sageexample}

@@

Besides basic arithmetic for Ore operators, the package provides a wide range of
methods to create, manipulate and solve several different kinds of operators.
Some of these methods are accessible in any Ore algebra while others are tied
specifically to, e.g., recurrence operators or differential operators.

In this section, we give an overview of the functionality provided by
the package. Because of space limitation, only some of the available methods
can be discussed here. For further information, we refer to the documentation.

@@

\begin{center}
  \begin{tabular}{|l|p{.65\hsize}|}
    \hline
    Method name & short description \\\hline
    \verb|A.quo_rem(B)| & returns the left quotient and the left remainder of $A$ and $B$.\\
    \verb|A.gcrd(B)| & returns the greatest common right divisor of $A$ and $B$.\\
    \verb|A.xgcrd(B)| & returns the greatest common right divisor of $A$ and $B$ and the according B\'ezout coefficients.\\
    \verb|A.lclm(B)| & returns the least common left multiple of $A$ and $B$.\\
    \verb|A.xlclm(B)| & returns the least common left multiple $L$ of $A$ and $B$ and the left quotients of $L$ and $A$ and of $L$ and $B$.\\
    \verb|A.resultant(B)| & returns the resultant of $A$ and $B$ (see \cite{li96} for its definition and
    properties).\\
    \hline
  \end{tabular}
\end{center}

@@

The GCRD is only unique up to multiplication (from the left) by an element from
the base ring. The method \verb|normalize| called in line~6 of the listing above
multiplies a given operator from the left by some element from the base ring
such as to produce a canonical representative from the class of all the operators
that can be obtained from each other by left multiplication of a base ring
element. This facilitates the comparison of output.

The efficiency of computing the GCRD depends on the size of the
coefficients of intermediate results, and there are different strategies to
control this growth via so called polynomial remainder sequences (PRS). The default is
the improved PRS described in~\cite{jaroschek13b}, which will usually be the
fastest choice. Other strategies can be selected by the option \verb|prs|.

@@

If $L_1,L_2$ are operators, then the solutions of their GCRD are precisely the
common solutions of $L_1$ and~$L_2$. The LCLM, on the other hand, is the minimal
order operator whose solution space contains all the solutions of $L_1$ and
all the solutions of~$L_1$. Because of this property, GCRD and LCLM are
useful tools for constructing operators with prescribed solutions. For example,
here is how to construct a differential operator which has the solutions $x^5$
and $\exp(x)$, starting from the obvious operators annihilating $x^5$
and~$\exp(x)$, respectively.

@@

Thanks to the LCLM operation discussed above, we have the property that when $f$
and $g$ are two objects which are annihilated by some operators $L_1,L_2$
belonging to some Ore algebra~$A$ then this algebra contains also an operator
which annihilates their sum~$f+g$. In other words, the class of solutions of operators of
$A$ is closed under addition. It turns out that similar closure properties hold
for other operations. The following table lists some of the corresponding
methods. Methods for more special closure properties will appear further below.

@@

As an example application, let us prove Cassini's identity for Fibonacci
numbers: 
\[
  F_{n+1}^2 - F_nF_{n+2}=(-1)^n.
\]
The idea is to derive, using commands from the table above, a recurrence
satisfied by the left hand side, and then show that this recurrence is also
valid for the right hand side.

@@

For the elements of some of the most important algebras, additional methods have
been implemented. The following table lists some of the additional methods
available for differential operators, i.e., elements of an Ore algebra of the
form~$R[x]\<Dx>$ or $K(x)\<Dx>$.

@@

The last output implies that the operator annihilating $\int_0^x \exp(-t^2) dt$
also admits a solution which behaves for $x\to\infty$ like $\frac1x\exp(-x^2)$.

The next example illustrates the methods for finding rational and polynomial
solutions of an operator~$L$.  These methods accept as an optional parameter an
inhomogeneous part consisting of a list (or tuple) of base ring elements,
$(f_1,\dots,f_r)$. They return as output a list of tuples $(g,c_1,\dots,c_r)$
with $L(g)=c_1f_1+\cdots+c_rf_r$ where $g$ is a polynomial or rational function
and $c_1,\dots,c_r$ are constants, i.e., elements of the base ring's base
ring. The tuples form a vector space basis of the solution space.

In the example session below, we start from two polynomials $p,q$, then compute an 
operator~$L$ having $p$ and $q$ as solutions, and then recover $p$ and $q$ from~$L$.
Note that for consistency also the solutions of homogeneous equations are returned as 
tuples. At the end we give an example for solving an inhomogeneous equation.

@@

In some cases, for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$, isolated values of a sequence can be
computed asymptotically faster for large $n$ than by listing all
values, using the binary splitting technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix $P$ and a polynomial $Q$ such that $P / Q$ multiplied by a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $e$ takes a fraction of a second:

@@

Guessing is, in some sense, the reverse operation of \verb|to_list| for
recurrence operators, or of \verb|power_series_solutions| for differential
operators.  It is of the most popular features of packages like gfun, and there
are even some special purpose packages dedicated to this
technique~\cite{kauers09a,hebisch11}. The basic idea is simple. Given
a finite array of numbers, thought of as the first terms of an infinite
sequence, we want to know whether this sequence satisfies a recurrence. The
algorithm behind a guessing engine searches for small equations matching the
given data.  Generically, no such equations exist, so if some are found, it is
fair to ``guess'' that they are in fact valid equations for the whole infinite
sequence.

@@

Several options are available for customizing the search for relations.  In
order to explain them, we first need to give some details on the underlying
algorithms. For simplicity of language, we restrict here to the case of
recurrence operators. The situation for differential operators is very similar.

@@

\hangindent=-6.3cm\hangafter=-13\leavevmode
\smash{\raise-5.3cm\rlap{\kern\hsize\kern-5.1cm\begin{tikzpicture}[scale=.45]
  \draw[->] (10.5,0) node {$r$} (-.2,0) -- (10, 0) ;
  \draw[->] (0,10.5) node {$d$} (0,-.2) -- (0, 10) ;
  \draw (1.2, 3.5) -- (1.4, 4) (1.2, 3) -- (1.6, 4) (1.2, 2.5) -- (1.8, 4) (1.2, 2) -- (2, 4)
        (1.4, 2) -- (2.2, 4) (1.6, 2) -- (2.4, 4) (1.8, 2) -- (2.6, 4) (2, 2) -- (2.8, 4) 
        (2.2, 2) -- (3, 4) (2.4, 2) -- (3.2, 4) (2.6, 2) -- (3.4, 4) (2.8,2)--(3.6,4) (3,2)--(3.8,4)
        (3.2, 2) -- (4, 4) (3.4, 2) -- (4.2, 4) (3.6, 2) -- (4.4, 4) (3.8,2)--(4.6,4) (4,2)--(4.8,4)
        (4.2, 2) -- (5, 4) (4.4, 2) -- (5.2, 4) (4.6, 2) -- (5.4, 4) (4.8,2)--(5.6,4) (5,2)--(5.8,4); 
  \fill[lightgray] (.75, 10) .. controls (1.5,2.5) and (2.5,1.5) .. (10, .75) -- (10, 10) -- (.75, 10);
  \draw[thick] (2, 10) .. controls (2, 1.2) and (2, 1.2) .. (10, 1.2); 
  \draw (-.2, 2) node {\vbox{\llap{\footnotesize min }\kern-3pt\llap{\footnotesize degree }}} -- (10, 2) 
        (-.2, 4) node {\vbox{\llap{\footnotesize max }\kern-3pt\llap{\footnotesize degree }}} -- (10, 4);
  \draw (1.2, -.2) node {\hbox to0pt{\rule{0pt}{2em}\hss\footnotesize min order\hss}} -- (1.2, 10)
        (5, -.2) node {\hbox to0pt{\rule{0pt}{2em}\hss\footnotesize max order\hss}} -- (5, 10);
  \fill (2.5, 3) circle(4pt) node[right] {\rule[-1em]{0pt}{0pt}$A$} (2, 9) circle(4pt) node[right] {$B$} ;
\end{tikzpicture}}}%
The figure on the right illustrates the typical situation for guessing problems
that are not too small and not too artificial. The gray region indicates
the area which is not accessible with the given amount of data. Only the points
$(r,d)$ below it can be tested for an operator of order~$r$ and degree~$d$ that
fits to the given data. Let's assume that operators exist on and above the solid
black hyperbola. The user will usually not know this curve in advance but may
have some expectations about it and can restrict the search accordingly, for
example to the dashed area shown in the figure. The method will detect the
existence of an operator, say at point~$A$, and construct from the information
gained at this point an operator of minimal possible order, which may correspond
to point~$B$.  This operator is returned as output. Note that the degree of the
output may exceed the value of \verb|max_degree|, and its order may be smaller
than $\verb|min_order|$:

@@

In order to test a specific point $(r,d)$, the data array must contain at least $(r+1)(d+2)$ terms. 
If it has more terms, the guess becomes more trustworthy, but also the computation time increases.
By setting the option \verb|ensure| to a positive integer~$e$, the user can request that only 
such points $(r,d)$ should be tested for which the data array contains at least $e$ more terms than
needed. This increases the reliability. By setting the option \verb|cut| to a positive integer~$c$,
the user requests that for testing a point $(r,d)$, the method should take into account at most $c$
more terms than needed. If the data array contains more terms, superfluous ones are ignored in the
interest of better performance. We must always have $0\leq e\leq c\leq \infty$. The default setting
is $e=0$, $c=\infty$.
